---
title: "DATASHEET: Nitrate Vulnerable Zone (NVZ) designation 2017 - Surface Water"
subtitle: "NVZ ID: `r nvz.id` NVZ Name: `r nvz.name`"
date: "Publication Date: `r date`"
output: 
    pdf_document: 
      fig_caption: yes
      latex_engine: xelatex
      number_sections: yes
      toc_depth: 4
mainfont: Arial
fontsize: 11pt
linkcolor: black
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyfoot[CO,CE]{NVZ ID; `r nvz.id` - NVZ Name; `r nvz.name`}
- \fancyfoot[LE,RO]{\thepage}
---

```{r library , echo = FALSE}
library(knitr)
library(markdown)
library(rmarkdown)
library(png)
library(grid)
library(ggplot2)
```


```{r front_image, echo = FALSE, fig.width = 7, fig.height = 7, results = "asis"}

img <- readPNG("front_image.png")
plot.new()
grid.raster(img)
```

\newpage

This document contains a summary of the evidence used to designate NVZ `r nvz.id` - `r nvz.name.short`. It was produced on `r date` by the Environment Agency on behalf of the Secretary of State for the Environment.

`r additional.info` 

`r nvz.name.short` is `r nvz.type1.text` designation. There `r fail.count.text` in the waterbody that designates this NVZ, and, in total, `r fail.count.text.upstream` in the entire NVZ including the designating waterbody and any area upstream of the designating waterbody. `r mon.change.text.long` `r mod.change.text.long` `r nvz.des.text`  

The contribution of agriculture to the observed nitrate pollution in this NVZ has been assessed and we are confident that agriculture is an important contributor to nitrate pollution.

\clearpage

\tableofcontents

\newpage

# Introduction

**NVZ name: `r nvz.name.short`**  
**NVZ ID: `r nvz.id`** 

This document provides a summary of the evidence used in proposing an area of land as one which should be, or should continue to be, designated as a Nitrate Vulnerable Zone (NVZ) for the purposes of the Nitrate Pollution Prevention Regulations 2015. 

A full description of the methods used in developing the NVZ proposals is set out in the detailed methodology for surface water NVZs, available via [http://apps.environment-agency.gov.uk/wiyby/141443.aspx](apps.environment-agency.gov.uk/wiyby/141443.aspx) . These methods were developed under the guidance of a review group convened by the Defra for the last NVZ review (2011-2013), which included representatives from the farming and water industries as well as independent academic experts. Minor refinements to the methods have been made for the current review.

Under the surface water method a water is designated if is it affected by pollution or could be affected by pollution.

---

Please note that any maps shown here have not used detailed field boundaries and therefore represent the indicative 'soft' boundary only. The definitive NVZ area can be seen on the “What’s in Your Backyard”(WIYBY) website [http://apps.environment-agency.gov.uk/wiyby/141443.aspx](apps.environment-agency.gov.uk/wiyby/141443.aspx).


The maps featured in this report are subject to at least one of the following copyright statements;  
&copy; Crown copyright and database rights 2016 Ordnance Survey 100024198.   
&copy; Environment Agency copyright and/or database rights 2016. All rights reserved.  
&copy; Centre for Ecology and Hydrology 2016.  

\newpage

## Core principles  

1. Each NVZ review takes the previous designations as the starting point and most designations are continuations of those designations.
2. Evidence of pollution is based on a combination of monitoring data, recent and longer term trends, and modelled data.
3. A water is affected by pollution if our monitoring or modelling evidence, or a combination of the two, confidently assesses that the 95^th^ percentile[^1] concentration of nitrate exceeds 50 mgl^-1^ (as nitrate (NO~3~)).
4. A water could be affected by polltion if our monitoring evidence indicates 95^th^ percentile concentrations will exceed 50 mgl^-1^ nitrate in the near future (2020).
5. We include all inorganic nitrogen (N) compounds in our assessment, not just nitrate. 50 mgl^-1^ nitrate (as NO~3~) is equivalent to 11.3 mgl^-1^ nitrate (as N). As nitrogen can readily transform we include all commonly monitored forms of inorganic nitrogen, nitrate, nitrite and ammonium (referred to as Total Inorganic Nitrogen (TIN) when summed). All water quality data presented below is TIN and is compared to 11.3 mgl^-1^ to determine if the water is affected by pollution or could be affected by pollution.
6. If a sample point is downstream of a sewage treatment works or other point source consented discharge[^2] we assess whether the sample point is representative of the wider water quality of the catchment, in terms of nitrate. We do not automatically discount such sample points, a comprehensive national and local screening procedure is used to determine if a sample point should be removed from our analysis.
7. We use the Water Framework Directive river waterbodies as our units of assessment.  If one part of that waterbody is affected by pollution or could be affected by pollution then the entire waterbdoy is designated.  This is the case even if the sole monitoring point that shows evidence of pollution is in the upper (upstream) section of the waterbody and downstream monitoring points show no evidence of pollution. For 'catchment' designations the designated NVZ is all land that drains to the designating waterbody. For 'waterbody' designations the designated NVZ covers only the designating waterbody.
8. All existing designation are continued unless there we can be confident that the water is no longer affected by pollution and will not become affected by pollution. We have already considered whether existing designations are eligible for removal.
9. All proposed designations are developed close working between national and area Environment Agency staff with quality assurance through discussions at workshops.

We review the evidence every four years, and we believe that NVZ designation is justified in the area that is described in this data sheet. 

[^1]: A 95^th^ percentile is the value exceeded by 5% of samples. It is used instead of a maximum value as it more stable and less sensitive to extreme events.
[^2]: Consented discharge means all point source discharges that are consented to discharge to controlled waters by the Environment Agency. This includes sewage treatment works (sometimes also refereed to as waste water treatment works), industrial discharges, private sewage discharges, intermittent sewage discharges, landfills & all other discharge types.

\newpage



# Basic information

```{r tab_count_basic, echo = FALSE, results = "asis"}

tab.count <- TabCount(tab.count)
```

-------------------------------------------- ---------------------
NVZ
-------------------------------------------- ---------------------
NVZ Name                                     `r nvz.name.short`

NVZ ID                                       `r nvz.id`

Area (km^2^)                                 `r nvz.area`

NVZ type (existing/new)                      `r nvz.type1`

NVZ type (catchment/waterbody)               `r nvz.type2.text`

Designating WFD waterbody                    `r nvz.wb`

% of this NVZ previously designated as 
surface water NVZ                            `r nvz.prev.sw`

% of this NVZ previously designated as 
any NVZ                                      `r nvz.prev`
-------------------------------------------- ---------------------

: Designation summary

```{r fig_tab_count_matrix, echo = FALSE, results = "asis"}

fig.count <- FigCount(fig.count)
tab.count <- TabCount(tab.count)
```

Designations under the surface water method use a combination of monitored and modelled information.  Water quality data from our monitoring and modelling assessments are put into one of six classes, where enough data is available, that reflects how  confident we are that the 95^th^ percentile will exceed 11.3 mgl^-1^. When we combine the monitoring and modelled data, we use the following matrix (figure `r fig.count`) to identify those waters that are affected by pollution or could be affected by pollution. The combinations marked in red indicate a new designation should be recommended where there is not an existing NVZ designation. Where there is already an NVZ designation, the designation is continued unless there is sufficient evidence for de-designation. 

```{r matrix, echo = FALSE, fig.height = 3, fig.width = 3, results = "asis", fig.cap = "Evidence matrix used to determine if waterbody is subject to nitrate pollution"}

img <- readPNG("matrix.png")
plot.new()
grid.raster(img)
```

The 1 to 6 classes used in the matrix are explained in table `r tab.count`. A class of zero is given where there are no suitable sample points within the waterbody, so a monitoring class can't be given, or where the waterbody was not large enough to be used in the modelling assessment.

\newpage

--------- -------------------------------------------------
Class     Description
--------- -------------------------------------------------
1         High confidence 95^th^ percentile concentration is **below** 11.3 mgl^-1^

2         Moderate confidence 95^th^ percentile concentration is **below** 11.3 mgl^-1^

3         Low confidence 95^th^ percentile concentration is **below** 11.3 mgl^-1^

4         Low confidence 95^th^ percentile concentration is *above* 11.3 mgl^-1^

5         Moderate confidence 95^th^ percentile concentration is *above* 11.3 mgl^-1^

6         High confidence 95^th^ percentile concentration is *above* 11.3 mgl^-1^

0         Monitoring or modelling classification not available 
--------- -------------------------------------------------

: Matrix class descriptions

```{r tab_count_scores, echo = FALSE, results = "asis"}

tab.count <- TabCount(tab.count)
```

The designating waterbody has had the following `r main.wb.scores.text` in the last three NVZ reviews (table `r tab.count`);

```{r basic_scores, echo = FALSE, results = "asis"}
mon.scores <- c(mon.2009, mon.2013, mon.2017)
mod.scores <- c(mod.2009, mod.2013, mod.2017)
scores.tab <- data.frame(mon.scores, mod.scores, row.names = c(2009, 2013, 2017))
names(scores.tab) <- c("Monitoring scores", "Modelling scores")
kable(scores.tab, caption = "Matrix scores for the last three NVZ reviews")
```


`r nvz.name.short` is `r nvz.type1.text` designation.  `r mon.change.text.long` `r mod.change.text.long` `r nvz.des.text`

```{r figcount_overviewmap, echo = FALSE, results = "asis"}

fig.count <- FigCount(fig.count)
```

Figure `r fig.count` shows the area covered by this NVZ.  

*	The pink outline indicates the extent of this NVZ.
*	The yellow area shows any surrounding area covered by other surface water NVZs.
*	The black hatched area shows the designating waterbody that is either affected by pollution or could be affected by pollution.


![Overview of designated catchment](`r overview.map`)  


\newpage 

# Monitoring Evidence

```{r mon_main, echo = FALSE, results = "asis"}

tab.count <- TabCount(tab.count)

mon.yes <- FALSE
mon.pholder.text <- "Monitoring evidence was not used in this designation. There were not any monitoing points with sufficent data in the designation waterbody. Table XX will contain NA (Not Available) values"

if (mon.2017 != 0) {
    mon.yes <- TRUE
}

if (!mon.yes) {
    cat("\n", mon.pholder.text, "\n")
}
```

------------------------------------------------------- --------------------------------
NVZ
------------------------------------------------------- --------------------------------
Number of sample points in the designating
waterbody                                               `r smpt.count`

Number of sample points in the entire designated NVZ    `r smpt.count.upstream`

Number of sample points with 95^th^ percentile TIN 
greater than 11.3 mgl^-1^ in the designating waterbody  `r fail.count`

Number of sample points with 95^th^ percentile TIN 
greater than 11.3 mgl^-1^ in the entire designated NVZ  `r fail.count.upstream`

Most polluted sample point in the designating 
waterbody                                               `r worst`

Current monitored 95^th^ percentile 
concentration (class)                                   `r paste0(mon.conc, " mgl^-1^ (", mon.2017, ")")`

High confidence lower and upper bounds for current 
monitored 95^th^ percentile concentration               `r mon.conc.l95` mgl^-1^ - `r mon.conc.u95`  mgl^-1^

Moderate confidence lower and upper bounds for current 
monitored 95^th^ percentile concentration               `r mon.conc.l75 `  mgl^-1^ - `r mon.conc.u75`  mgl^-1^

Trend (future) 95^th^ 
percentile concentration (class)                        `r paste0(trend.conc, " mgl^-1^ (", trend.2017, ")")`

High confidence lower and upper bounds for trend 
95^th^ percentile concentration                         `r trend.conc.l95`  mgl^-1^ - `r trend.conc.u95`  mgl^-1^

Moderate confidence lower and upper bounds for trend
95^th^ percentile concentration                         `r trend.conc.l75`  mgl^-1^ - `r trend.conc.u75`  mgl^-1^

------------------------------------------------------- --------------------------------

: Summary of monitoring evidence

There `r fail.count.text` in the waterbody that designates this NVZ, and, in total, `r fail.count.text.upstream` in the entire NVZ including the designating waterbody and any area upstream of the designating waterbody.

```{r mon_plot, echo = FALSE, results = "asis", fig.width = 6, fig.height = 6, fig.cap = "TIN as N time series for worst sample point in designating waterbody"}
if (mon.yes) {
    img <- readPNG(worst.graph)
    plot.new()
    grid.raster(img)
}
```

```{r mon_text1, echo = FALSE, results = "asis"}
if (mon.yes) {
    fig.count <- FigCount(fig.count)
    mon.text1 <- paste0("Figure ", fig.count, " shows the time series of concentrations for the most polluted sample point in the designating waterbody. In figure ", fig.count, "; the circles represent each sample taken since 1990; the dark blue line represents the 95^th^ percentile over time; ", curr.est.text, "; the right most dashed blue line represents mid-2020, where it crosses the dark blue line is the future (2020) 95^th^ percentile; the blue shaded areas from 2015 onward represent the uncertainty in the future 95^th^ percentile, we have high confidence that the future 95^th^ percentile lies within the light blue shaded area and moderate confidence that the future 95^th^ percentile lies within the dark blue shaded area.")
    cat("\n", mon.text1, "\n")
}
```

```{r mon_text2, echo = FALSE, results = "asis"}
if (smpt.count.upstream > 0) {
    mon.text2 <- "Graphs for all monitoring points in this NVZ designation are included in appendix 1."
    cat("\n", mon.text2, "\n")
}
```

```{r mon_text3, echo = FALSE, results = "asis"}
if (mon.yes) {
    fig.count <- FigCount(fig.count)
    mon.text3 <- paste0("Figure ", fig.count, " shows the location of the monitoring points in the designating waterbody. The monitoring points are shown as coloured points, the colours show the monitoring class (1-6) of each point.  The monitoring score per sample point is based on the highest of the current and future 95^th^ percentiles. The monitoring score per catchment is taken from the most polluted sample point in the designating waterbody.")
    cat("\n", mon.text3, "\n")
}
```

In figure `r fig.count`, the small crosses represent the outlets of consented discharges[^2]. Any monitoring point that is too close (within the mixing zone) to a consented discharge of nitrogen compounds and therefore not representative of the water quality of the waterbody or is otherwise not representative of the water quality within the waterbody is excluded from the analysis.

![Overview of monitoring locations and classes with point source locations](`r monitoring.map`)

\newpage

# Modelling Evidence

Our modelling (referred to as the land use model) predicts the underlying risk of pollution from the land use within the designated NVZ catchment. It allows us to assess whether a waterbody is affected by pollution even if we have no monitoring data or where the monitoring data is not suitable for use in this assessment.

The land use model uses information on;  

*	Agricultural N losses
*	Consented discharges 
*	Diffuse urban N losses
*	Residential properties not connected to sewer networks

The water quality data used in the [monitoring evidence](#monitoring-evidence) chapter is also used in the development of the land use model.

``` {r land_main, echo = FALSE, results = "asis"}

tab.count <- TabCount(tab.count)

land.yes <- FALSE
land.pholder.text <- paste0("Outputs from the land use model were not used in this designation. This is because the area of the designation is less than  20 km^2^. Some of the input data to the land use model (the agricultural land use data) is only reliable at designation sizes grater than 20 km^2^. The land use model summary table (table ", tab.count, ") will contain NA (Not Available) values.")



if (mod.2017 != 0) {
    land.yes <- TRUE
}

if (!land.yes) {
    cat("\n", land.pholder.text, "\n")
}
```


-------------------------------------------------------------- ---------------------
NVZ
-------------------------------------------------------------- ---------------------
Land use model 95^th^ percentile 
concentration (class)                                          `r mod.conc` (`r mod.2017`)

High confidence lower and upper bounds for land use model 
95^th^ percentile concentration estimate                       `r mod.conc.l95` - `r mod.conc.u95`

Moderate confidence lower and upper bounds for land use model 
95^th^ percentile concentration estimate                       `r mod.conc.l75` - `r mod.conc.u75`

-------------------------------------------------------------- ---------------------

: Summary of modelling evidence



```{r land_plot, echo = FALSE, fig.width = 6, fig.height = 2, fig.cap = "Landuse analysis predictions relative to 11.3 mgl^-1^ TIN.", results = "asis"}
if (land.yes) {
    mod.concs.plot <- c(mod.conc.u95, mod.conc.u75, mod.conc, mod.conc.l75, 
                        mod.conc.l95)
    p <- LinesGraph(mod.concs.plot)
    p
}
```

```{r land_text1, echo = FALSE, results = "asis"}
if (land.yes) {
    fig.count <- FigCount(fig.count)
    land.text1 <- paste0("Figure ", fig.count, " shows the concentration of the 95^th^ percentile TIN concentration land use model prediction, plus the upper and lower, moderate and high confidence bounds in relation to 11.3 mgl^-1^ TIN.")
    cat("\n", land.text1, "\n")
}
```

```{r fig_count_modelmap, echo = FALSE, results = "asis"}

fig.count <- FigCount(fig.count)
```

Figure `r fig.count` shows how the model predictions vary across the designated area. If any waterbodies are not coloured on figure `r fig.count` this is because that the waterbody or the upstream area of the waterbody has an area less than 20 km^2^. Figure `r fig.count` is included even if the land use model was not used in this designation.

![Landuse modelling predictions](`r modelling.map`)

\newpage

# Supporting Evidence

## Contribution of agriculture to nitrogen losses

### Sectoral load apportionment from land use analysis inputs

We consider a range of the most important nitrogen sources in a NVZ designation in our land use analysis.  Based on the data that is input to that analysis, we can estimate the contribution of each source to the total losses of nitrogen within the designated area.  


```{r load_main, echo = FALSE, results = "asis"}
load.yes <- FALSE
load.pholder.text <- "Load apportionment estimates are not available for this NVZ designation."

if (nvz.area > 20) {
    load.yes <- TRUE
}

if (load.yes) {
        tab.count <- TabCount(tab.count)
    load.text1 <- paste0("We present two figures for load from consented discharges[^2]. The maximum value is based on discharge consent conditions (the conditions consented discharges may not exceed). Therefore the maximum value represents the load emitted from a combination of the greatest flow and the greatest concentration of nitrogen compounds all consented discharges in the catchment may legally emit. The minimum value is based on information from consented discharges with flow and nitrogen concentration monitoring, so represents the most likely load from all consented discharges in the catchment.

Table ", tab.count, " shows estimated load across the designated area by individual sectors. The minimum agricultural contribution is based on the maximum consented discharge value, while the maximum agricultural contribution uses minimum consented load. For the agricultural load, diffuse urban load and the unsewered residential properties load there is a single value only.")
    cat("\n", load.text1, "\n")
} else{ cat("\n", load.pholder.text, "\n")}
```

```{r load_table, echo=FALSE, results="asis"}
if (load.yes) {

    minag.loads <- c(ps.load.max, diff.urban.load, septic.load, ag.load,
                     total.load.agmin, perc.ag.load.min)
    maxag.loads <- c(ps.load.min, diff.urban.load, septic.load, ag.load,
                     total.load.agmax, perc.ag.load.max)
    load.names <- c("Consented discharges (kg y^-1^)", "Diffuse urban (kg y^-1^)", 
                    "Unsewered residential properties (kg y^-1^)", 
                    "Agriculture (kg y^-1^)", "Total cathcment load (kg y^-1^)", 
                    "% of total load from agriculture")
    loads.table <- data.frame(minag.loads, maxag.loads, row.names = load.names)
    names(loads.table) <- c("Miniumum agricultural contribution", "Maximum agricultural contribution") 
    kable(loads.table, caption = "Summary of sources of pollution by nitrogen 
      compounds")
}
```

```{r load_text2, echo=FALSE, results = "asis"}

if (load.yes) {
    fig.count  <- FigCount(fig.count)
    fig.count2 <- FigCount(fig.count)
    load.text2 <- paste0("Figures ", fig.count, " and ", fig.count2, " shows the information in table ", tab.count, " presented as percentage apportionment to individual sectors. Figure ", fig.count, " shows the minimum agricultural contribution and figure ", fig.count2, " shows the maximum agricultural contribution.")
    cat("\n", load.text2, "\n")
}
```

``` {r load_plot1, echo = FALSE, fig.width = 6, fig.height = 2, fig.cap = "Percentage apportionment to sectors using the minimum agricultural load estimate and the maximum consented discharge load estimate from 2017 NVZ designation land use analysis"}

if (load.yes) {
    min.la.perc <- ((loads.table[1:4, 1]) / loads.table[5, 1]) * 100
    la.labs <- c("Consented discharges", "Diffuse urban", 
                 "Unsewered residential properties", "Agriculture")
    min.la.perc <- data.frame(variable = la.labs, value = min.la.perc)
    
    p <- ggplot(min.la.perc, aes(x = reorder(variable, value), y = value))
    p <- p + geom_bar(stat = "identity", fill = "grey60") + coord_flip()
    p <- p + geom_text(aes(label = paste0(round(value, 1), "%")), size = 3)
    p <- p + ylim(0, 100)
    p <- p + theme(panel.background = element_blank(),
                   axis.ticks = element_blank(),
                   axis.title = element_blank(),
                   axis.text.x = element_blank(),
                   legend.position = "none")
    p
}
```

``` {r load_plot2, echo = FALSE, fig.width = 6, fig.height = 2, fig.cap = "Percentage apportionment to sectors using the maximum agricultural load estimate and the minimum consented discharge load estimate from 2017 NVZ designation land use analysis"}

if (load.yes) {
    max.la.perc <- ((loads.table[1:4, 2]) / loads.table[5, 2]) * 100
    max.la.perc <- data.frame(variable = la.labs, value = max.la.perc)
    
    p <- ggplot(max.la.perc, aes(x = reorder(variable, value), y = value))
    p <- p + geom_bar(stat = "identity", fill = "grey60") + coord_flip()
    p <- p + geom_text(aes(label = paste0(round(value, 1), "%")), size = 3)
    p <- p + ylim(0, 100)
    p <- p + theme(panel.background = element_blank(),
                   axis.ticks = element_blank(),
                   axis.title = element_blank(),
                   axis.text.x = element_blank(),
                   legend.position = "none")
    p
}
```

Regardless of whether load apportionment estimates are available for this NVZ, a full list of discharges is included in appendix 2.

[^3]: Lee, D. et al., 2015. NEAP-N Nitrate leaching for 1970 and 2014, WT1550, Report to Defra.

\newpage

### Sectoral load apportionment from SEPERATE[^5] framework

SEPERATE (SEctor Pollutant AppoRtionment for the AquaTic Environment) is a national scale multiple pollutant source apportionment framework developed by ADAS, Rothamsted Research, Centre for Ecology and Hydrology and the Environment Agency. SEPERATE uses data on sources of total nitrogen including agricultural practice, sewage treatment works, combined sewer overflows, storm tanks & septic tanks, to apportion the load input to the aquatic environment into individual sectors.  

```{r seperate_text, echo=FALSE}




```

```{r seperate_main, echo = FALSE, results = "asis"}

sep.yes <- FALSE
sep.pholder.text <- "Outputs from the SEPERATE model are not available for this NVZ designation."

if (nvz.type3 == 2 && nvz.area >= 25) {
    sep.yes <- TRUE
    if (nvz.type2 == "C") {
        sep.dat <- sep_cumulative[sep_cumulative$EA_WB_ID == nvz.wb, ]
        sep.labs <- c("Agricultural", "Bank Erosion", "Urban Diffuse", "STW",
                      "Storm Tank", "Sceptic Tank", "CSO", "Direct Deposition", 
                      "Groundwater", "Waterbody Total")
    } 
    
    if (nvz.type2 == "WB") { 
        sep.dat <- sep_local[sep_local$EA_WB_ID == nvz.wb, ]
        sep.labs <- c("Agricultural", "Bank Erosion", "Urban Diffuse", "STW",
                  "Storm Tank", "Sceptic Tank", "CSO", "Direct Deposition",
                  "Waterbody Total")
    }
}

if (nvz.type3 == 1 && nvz.area >= 25) {
    if (nvz.type2 == "C" && nvz.wb %in% sep_cycle1.us$EA_WB_ID) {
        sep.yes <- TRUE
        sep.dat <- sep_cumulative[sep_cumulative$EA_WB_ID == nvz.wb, ]
        sep.labs <- c("Agricultural", "Bank Erosion", "Urban Diffuse", "STW",
                      "Storm Tank", "Sceptic Tank", "CSO", "Direct Deposition", 
                      "Groundwater", "Waterbody Total")
    } 
    
    if (nvz.type2 == "WB" && nvz.wb %in% sep_cycle1$EA_WB_ID) {
        sep.yes <- TRUE
        sep.dat <- sep_local[sep_local$EA_WB_ID == nvz.wb, ]
        sep.labs <- c("Agricultural", "Bank Erosion", "Urban Diffuse", "STW",
                  "Storm Tank", "Sceptic Tank", "CSO", "Direct Deposition",
                  "Waterbody Total")
    }
}

if (sep.yes) {
    tab.count <- TabCount(tab.count)
    fig.count2 <- FigCount(fig.count2)
    sep.labels.text <- paste0("Table ",  tab.count, " shows the load SEPERATE estimates is generated persector. Figure ", fig.count2, " shows the same data expressed as a percentage of the total load. Note that for main river designations data is provided for the entire upstream area, while for tributary designations only data from the designating waterbody is provided.")
    cat("\n", sep.labels.text, "\n")
} else {cat("\n", sep.pholder.text, "\n")}
```

```{r seperate_table, results="asis", echo=FALSE}
if (sep.yes) {
    sep.dat.load <- sep.dat[, 2:((ncol(sep.dat) / 2) + 1)]
    sep.dat.load <- data.frame(t(sep.dat.load))
    row.names(sep.dat.load) <- c(sep.labs)
    names(sep.dat.load) <- c("Load (t yr^-1^)")
    kable(sep.dat.load, caption = "Load in t yr^-1^ across designated area from SEPERATE framework (2014)")
}

```

```{r seperate_plot, echo = FALSE, fig.cap = "Percentage apportionment of nitrate load across designated area from SEPERATE framework (2014)", fig.height = 3, fig.width = 6}
if (sep.yes) {
    sep.dat.perc <- sep.dat[, c(((ncol(sep.dat) / 2) + 2):ncol(sep.dat))]
    sep.dat.perc <- cbind(sep.labs[-length(sep.labs)], data.frame(t(sep.dat.perc)))
    names(sep.dat.perc) <- c("variable", "value")
    p <- ggplot(sep.dat.perc, aes(x = reorder(variable, value), y = value))
    p <- p + geom_bar(stat = "identity", fill = "grey60") + coord_flip()
    p <- p + geom_text(aes(label = paste0(value, "%")), size = 3)
    p <- p + ylim(0, 100)
    p <- p + theme(panel.background = element_blank(),
                   axis.ticks = element_blank(),
                   axis.title = element_blank(),
                   axis.text.x = element_blank(),
                   legend.position = "none")
    p
}

```

[^5]: Zhang, Y. et al., 2014. Cross sector contributions to river pollution in England and Wales: Updating waterbody scale information to support policy delivery for the Water Framework Directive. Environmental Science and Policy, 42, pp.16-32.

\newpage

### Sectoral load apportionment from SIMCAT-SAGIS

SIMCAT (SIMulated CATchment) is a well-established river water quality model with a long history of use in the UK. It has recently been incorporated into a source apportionment framework, SAGIS (Source Apportionment Geographical Information System), to allow apportionment of a range of pollutants to various sectors, including agriculture. For catchment designations we provide the nitrate as N concentration SIMCAT predicts at the farthest downstream point of the designation and the proportion of that concentration from diffuse sources[^6].

```{r simcat, echo = FALSE, resutls = "asis"}
if (exists("simcat.conc")) {
    if (!is.na(simcat.conc)) {
        tab.count <- TabCount(tab.count)
        simcat <- c(simcat.conc, simcat.prop * 100)
        simcat <- data.frame(Result = simcat)
        row.names(simcat) <- c("SIMCAT prediction nitrate as N concentration at outlet of designated area (mgl^-1^)", "Percentage of concentration predicted to be from diffuse source (%)")
        kable(simcat, caption = "SIMCAT-SAGIS predictions for this designation")
        
    } else {cat("SIMCAT-SAGIS predictions are not avialable for this designation")
    }
} else {cat("SIMCAT-SAGIS predictions are not avialable for this designation")
}
```

[^6]: In SIMCAT-SAGIS diffuse sources include agriculture, but not solely agriculture. We provide the proportion from diffuse source figure to give context, and hopefully, corroboration to the other load apportionment methods presented above.




\newpage

## Land cover & livestock

We have used land cover and livestock information as part of the method to derive this designation. It is used to drive the NEAP N model[^3] and the estimate of urban N load which are both input to the land use model. We take our land cover and livestock data from the 2014 Defra June agricultural survey[^7]. We have included the data below for this designation.

```{r lc_plot, echo = FALSE, results = "asis", fig.cap = "Landcover area by category from 2014 Defra agricultural survey", fig.height = 3, fig.width = 6}
fig.count2 <- FigCount(fig.count2)
landuse <- c(arable.area, grass.area, rough.area, wood.area, 
             urban.area)
landuse.names <- c("Arable", "Permanent grassland", "Rough grazing",
                   "Woodland","Urban")
landuse <- data.frame(name = landuse.names, Area = landuse)
p <- ggplot(landuse, aes(x = reorder(name, Area), y = Area))
p <- p + geom_bar(stat = "identity", fill = "grey60") + coord_flip()
p <- p + geom_text(aes(label = paste(Area, "(Ha)")), size = 3)
p <- p + ylim(-25, max(landuse$Area * 1.1))
p <- p + theme(panel.background = element_blank(),
               axis.ticks = element_blank(),
               axis.title = element_blank(),
               axis.text.x = element_blank(),
               legend.position = "none")
p

```

```{r livestock_table, echo = FALSE, results = "asis"}
tab.count <- TabCount(tab.count)
livestock <- c(count.cows, count.sheep, count.pigs, 
               count.poultry, count.other)
livestock <- data.frame(Count = livestock)
row.names(livestock) <- c("Cattle", "Sheep", "Pigs", "Polutry",
                          "Other livestock")
kable(livestock, caption = "Count of livestock within NVZ designation from 2014 Defra agricultural survey")
```

[^7]: Defra, 2015. Agriculture in the United Kingdom 2014. HMSO. 

\newpage

## Agricultural load over time

```{r fig_count_neap, echo = FALSE, results = "asis"}
fig.count2 <- FigCount(fig.count2)
```

Figure `r fig.count2` shows how the nitrate available for leaching from agricultural sources, as predicted by the ADAS NEAP N model[^3], varies over time for this NVZ designation. The data is presented relative to a baseline (the year 2000). The NEAP N model uses long term average climate data, so changes in nitrate available for leaching is due solely to changes in cropping and livestock within the NVZ. 

![Nitrate avaialble for leaching across NVZ `r nvz.id` relative to 2000 baseline due to cropping and livestock](`r neap.graph`)

NEAP N has been developed over time so a portion of any temporal variation may be due to changes to the model, model parameters or input data. 

\newpage 

## Workshop discussion

As part of the designation process the Environment Agency held a series of workshops with EA area staff and invited partners, including representatives from the farming industry.

The purpose the workshops was to discuss any proposed NVZ designations where the evidence supporting designation was uncertain or contradictory, or where there were new or modified designations. We were not able to discuss every designation, but where discussion did take place abridged notes are provided below.

`r workshop.discussion`

\newpage  

# Appendices

## Appendix 1 - Sample point time series graphs

This page has been left intetionally blank.

\newpage

```{r all_smpt_plots, echo = FALSE, fig.width = 6, fig.height = 4.1, results = "asis", fig.keep = "all"}
for (i in 1:length(mult.graph)){
    cat("\n Figure A", i, ": ", all.smpts[i], " - ", all.smpts.names[i], sep = "")
    cat("\n")
    cat("\n")
    img <- readPNG(mult.graph[i])
    plot.new()
    grid.raster(img)
    cat("\n")
}
```

\newpage  

## Appendix 2 - Discharge constent conditions

``` {r dischrages_table, echo = FALSE, results = "asis"}
kable(source.table, caption = paste0("Summary of sources of nitrogen 
                                     compounds within NVZ", nvz.id))
```

In the table above flow is given in mega  per day (Mld^-1^), concentration in milligrams per litre (mgl^-1^ as N) expressed as N and load in kilograms per day as N (kgd^-1^ as N). 

[^8]: Lerner, D., 2000. Guidelines for estimating urban loads of nitrogen to groundwater, NT1845, Report to Defra.

\newpage

End of document

\newpage

```{r back_image, echo = FALSE, fig.width = 7, fig.height = 10, results = "asis"}
img <- readPNG("back_image.png")
plot.new()
grid.raster(img)
```
